---
title: "Code for Angio-AP analysis"
author: "García-Mulero S"
date: "12/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries, genesets, data

Load libraries
```{r}
require(reshape); require(ggplot2); library(ggpubr);  
require(RColorBrewer); library(GSVA); 
library(survival); library(ggthemes); library(survminer)
```

Load data
```{r}
load("data/clin-ex-info_ALL.Rdata")

```

Load pathways
```{r}
# Load broadSet form GSEABase package
load("data/Pathways.Rdata")
Pathways

genesets <- list(Angiogenesis=Pathways$BIOCARTA_VEGF_PATHWAY,
  Antigen_Presentation=Pathways$REACTOME_ANTIGEN_PRESENTATION)
```

Genesets enrichment
```{r}
GSVA.TCGA <- gsva(expr=as.matrix(exTCGA), gset.idx.list=genesets)
GSVA1 <- gsva(expr=as.matrix(ex1), gset.idx.list=genesets)
GSVA2 <- gsva(expr=as.matrix(ex2), gset.idx.list=genesets)
GSVA3 <- gsva(expr=as.matrix(ex3), gset.idx.list=genesets)
GSVA4 <- gsva(expr=as.matrix(ex4), gset.idx.list=genesets)

# join only GSVA results for pca
GSVA <- cbind(GSVA.TCGA, GSVA1, GSVA2, GSVA3, GSVA4)

data.st <- as.matrix(t(GSVA))
intersect(as.character(rownames(data.st)), as.character(dat$ID))

dat <- dat[order(dat$ID),]; data.st <- data.st[order(rownames(data.st)),]
identical(as.character(rownames(data.st)), as.character(dat$ID))

```


Create groups
```{r}

## perform high/ low labels separatelly in each dataset
## the, join datasets

# indicate 0 as cutoff
# combine variables

datTCGA <- merge(clinTCGA, t(GSVA.TCGA), by.x="ID", by.y="row.names")
datTCGA$Angiogenesis.Q <- ifelse(datTCGA$Angiogenesis>=0, "High Angio", "Low Angio")
datTCGA$Antigen_Presentation.Q <- ifelse(datTCGA$Antigen_Presentation>=0, "High AP", "Low AP")
datTCGA$Combined <- paste0(datTCGA$Angiogenesis.Q, " - " , datTCGA$Antigen_Presentation.Q)

dat1 <- merge(clin1, t(GSVA1), by.x="ID", by.y="row.names")
dat1$Angiogenesis.Q <- ifelse(dat1$Angiogenesis>=0, "High Angio", "Low Angio")
dat1$Antigen_Presentation.Q <- ifelse(dat1$Antigen_Presentation>=0, "High AP", "Low AP")
dat1$Combined <- paste0(dat1$Angiogenesis.Q, " - " , dat1$Antigen_Presentation.Q)

dat2 <- merge(clin2, t(GSVA2), by.x="ID", by.y="row.names")
dat2$Angiogenesis.Q <- ifelse(dat2$Angiogenesis>=0, "High Angio", "Low Angio")
dat2$Antigen_Presentation.Q <- ifelse(dat2$Antigen_Presentation>=0, "High AP", "Low AP")
dat2$Combined <- paste0(dat2$Angiogenesis.Q, " - " , dat2$Antigen_Presentation.Q)

dat3 <- merge(clin3, t(GSVA3), by.x="ID", by.y="row.names")
dat3$Angiogenesis.Q <- ifelse(dat3$Angiogenesis>=0, "High Angio", "Low Angio")
dat3$Antigen_Presentation.Q <- ifelse(dat3$Antigen_Presentation>=0, "High AP", "Low AP")
dat3$Combined <- paste0(dat3$Angiogenesis.Q, " - " , dat3$Antigen_Presentation.Q)

dat4 <- merge(clin4, t(GSVA4), by.x="ID", by.y="row.names")
dat4$Angiogenesis.Q <- ifelse(dat4$Angiogenesis>=0, "High Angio", "Low Angio")
dat4$Antigen_Presentation.Q <- ifelse(dat4$Antigen_Presentation>=0, "High AP", "Low AP")
dat4$Combined <- paste0(dat4$Angiogenesis.Q, " - " , dat4$Antigen_Presentation.Q)


# join all clinical
dat <- rbind(datTCGA[,c(1, 26:27, 41, 34, 42:46)],
	dat1[, c(1, 19:20, 22, 23:28)],
	dat2[, c(1, 17:18, 20, 21:26)],
	dat3[, c(1, 12:13, 11, 14:19)],
	dat4[, c(1, 10, 13, 12, 14:19)]) ## 213 samples

dat$Recurrence <- ifelse(dat$DFS==1, "Recurrent", "Non-recurrent")
dat$"Angiogenesis-Antigen Presentation" <- dat$Combined

# save in an RData
save(datTCGA, exTCGA, dat1, ex1, dat2, ex2, dat3, ex3, dat4, ex4, dat, ex_comb,
  file="data/angio-ap-info_ALL.Rdata")
```


Survival analysis
```{r}
# fit survival function
coxDFS <- coxph(Surv(as.numeric(DFS.Months), as.numeric(DFS)) ~ dat$"Angiogenesis-Antigen Presentation" + strata(dataset), data=dat)
fitDFS <- survfit(Surv(as.numeric(DFS.Months), as.numeric(DFS)) ~ dat$"Angiogenesis-Antigen Presentation", data=dat) # for dfs

kmDFS <- ggsurvplot(fitDFS, data=dat, 
            font.x = c(10, "plain", "black"),
            font.y = c(10, "plain", "black"),
            font.title= c(10, "plain", "black"),
            pval = TRUE, pval.size=4, pval.coord = c(10, 0.1),
            font.legend = c(9, "plain", "black"), 
            palette=c("firebrick3", "lightpink1","skyblue1", "dodgerblue4"), 
            ylab="Progression free survival probability",
            legend.title = "",
			      legend.labs = c("High Angio - High AP", "High Angio - Low AP", "Low Angio - High AP", "Low Angio - Low AP"),
            legend = c(0.75, 0.9)) 


png("res/KM_combined_ALL_broadsets.tiff", 
   width = 4.5, height = 4, units = "in" ,  res = 400);
  kmDFS$plot
dev.off()

### 5-years relapse rate (5 years == 60 months)

rate_HH <- survfit(Surv(as.numeric(DFS.Months), as.numeric(DFS))~1, data=dat[dat$Combined=="High Angio - High AP",])
summary(rate_HH)

rate_LL <- survfit(Surv(as.numeric(DFS.Months), as.numeric(DFS))~1, data=dat[dat$Combined=="Low Angio - Low AP",])
summary(rate_LL)
```

Angio-AP correlation
```{r}
head(dat)

# density plots
dat.m <- melt(dat[, c("Angiogenesis", "Antigen_Presentation")])

density <- ggplot(aes(x=value, color=variable), data=dat.m) +
	  geom_density() + xlim(-1,1)

jpeg("res/density_scores_ALL.jpg", 
 height=5, width=6,  res = 600, units="in")      
density 
dev.off()


# dotplot con 0  y los valores -1 a 1 y correlation lines
graph <- ggplot(dat) +
  geom_point(aes(x=Antigen_Presentation, y=Angiogenesis, color=Recurrence), size=3) +
  scale_color_manual(values=c("grey75", "grey35")) +
  geom_hline(yintercept= 0, linetype="dashed", color = "red") +
  geom_vline(xintercept = 0, linetype="dashed", color = "red") +
  geom_smooth(aes(x=Antigen_Presentation, y=Angiogenesis, color=Recurrence), method=lm, se=FALSE, fullrange=TRUE) +
  geom_smooth(aes(x=Antigen_Presentation, y=Angiogenesis), color="black", method=lm, se=FALSE, fullrange=TRUE) +
  scale_fill_manual(values = c("grey60", "grey40")) +
  stat_cor(aes(x=Antigen_Presentation, y=Angiogenesis, color=Recurrence), label.x.npc=0.6, label.y.npc=0.05, hjust=0, method = "spearman") +
  stat_cor(aes(x=Antigen_Presentation, y=Angiogenesis), label.x.npc=0.6, label.y.npc=0.11, hjust=0, method = "spearman") +
  theme_classic() + xlab("Antigen Presentation") + border()
  #+geom_rug(aes(x=Antigen_Presentation, y=Angiogenesis, color=Recurrence))

jpeg("res/dotplot_scores_ALL_Correlation.jpg", 
 height=4.5, width=6,  res = 600, units="in")      
graph 
dev.off()

```


Barplot of recurrence by group
```{r}
table(dat$Combined, dat$Recurrence)

addmargins(table(dat$Combined, dat$Recurrence))
perc_table <- addmargins(round(prop.table(table(dat$Combined, dat$Recurrence)),3))

# test
chisq.test(dat$Combined, dat$Recurrence)

### Plot percentages
cont_table <- as.matrix(table(dat$Combined, dat$Recurrence))
perc_table1 <- prop.table(cont_table,margin=1)*100
dat1 <- melt(perc_table1)
colnames(dat1) <- c("Combined", "Recurrence", "value")

# plots
barplot1 <- ggplot(data = dat1, aes(x = Combined, 
	y = value, fill = Recurrence)) + 
  geom_bar(stat = 'identity') +  
  theme_classic() +
  scale_fill_manual(values=c("grey85", "grey35")) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(y = "Frequency(%)", x="", title="Angiogenesis - Antigen Presentation") +
  theme(plot.title = element_text(size=10))
barplot1

jpeg("res/barplot1_angio-ap_ALL_DEF.jpg", 
     units="in", width=5, height=4, res=400)
barplot1
dev.off()
```

TIS score by group
```{r}
TIS <- c("CCL5", "CD27", "CD274", "CD276", "CD8A", "CMKLR1", "CXCL9", "HLA-DQA1", "HLA-DRB1", "HLA-E", "IDO1", "LAG3", "NKG7", "PDCD1LG2", "PSMB10", "STAT1", "TIGIT")

### TIS-scores between the four groups
GSVA.TCGA <- gsva(expr=as.matrix(exTCGA), gset.idx.list=list(TIS=TIS))
GSVA1 <- gsva(expr=as.matrix(ex1), gset.idx.list=list(TIS=TIS))
GSVA2 <- gsva(expr=as.matrix(ex2), gset.idx.list=list(TIS=TIS))
GSVA3 <- gsva(expr=as.matrix(ex3), gset.idx.list=list(TIS=TIS))
GSVA4 <- gsva(expr=as.matrix(ex4), gset.idx.list=list(TIS=TIS))

# join only GSVA results for pca
GSVA <- cbind(GSVA.TCGA, GSVA1, GSVA2, GSVA3, GSVA4)

# merge
dat_TIS <- merge(dat, t(GSVA), by.x="ID", by.y="row.names")

# plot

compare_means(TIS ~ Combined,  data = dat_TIS, method = "kruskal.test")
compare_means(TIS ~ Combined,  data = dat_TIS)
my_comparisons <- list( c("High Angio - High AP", "Low Angio - High AP"), c("High Angio - High AP", "High Angio - Low AP"))

TISplot <-ggplot(dat_TIS, aes(x=Combined, y=TIS)) +
         geom_boxplot(color="black", outlier.shape=NA) +
         geom_jitter(position=position_jitter(0.1), size=1.5, 
            aes(x=Combined, y=TIS, color=Recurrence)) +
         scale_color_manual(values=c("grey75", "grey35")) +
         theme_classic() + border() +
         theme(axis.text.x = element_text(angle=45, hjust=1)) +
         labs(x="", y = "TIS score") + annotate(geom="text", label="p < 2e-16", size=3, x=4, y=0.9)
         #stat_compare_means(method="kruskal.test", size=3, label.y.npc=0.95, label.x.npc=0.3, label="p.format")


png("res/TIS_score_ALL_combined.tiff", 
 height=3.5, width=5,  res = 600, units="in")      
TISplot
dev.off()
```

Months to reccurece by group
```{r}
### DFS between different groups
compare_means(DFS.Months ~ Combined,  data = dat[dat$DFS==1,], method = "kruskal.test")
compare_means(DFS.Months ~ Combined,  data = dat[dat$DFS==1,])

bp.cluster <-ggplot(dat[dat$DFS==1,], aes(x=Combined, y=DFS.Months)) +
         geom_boxplot(color="black", outlier.shape=NA) +
               geom_jitter(position=position_jitter(0.1), size=1.5, 
            aes(x=Combined, y=DFS.Months), alpha=0.5) +
         labs(x="", y = "Months to Recurrence") +
         theme(axis.text.x = element_text(angle=45, hjust=1)) +
         stat_compare_means(method="kruskal.test", 
          label="p.format", label.y.npc=0.95)

png("res/boxplot_DFS_groups_ALL.tiff", 
 height=3.5, width=4,  res = 600, units="in")      
bp.cluster
dev.off()
```

## Functional analysis

GSEA high-high VS low-low
```{r}
## create groups for GSEA
## only recurrent samples
dat2 <- dat[dat$Recurrence=="Recurrent",]
dat2 <- dat2[dat2$Combined=="High Angio - High AP" | dat2$Combined=="Low Angio - Low AP",]
table(dat2$Combined)

ex.gsva <- ex_comb[, colnames(ex_comb) %in% dat2$ID]
dim(ex.gsva)

ex.gsva <- ex.gsva[, order(colnames(ex.gsva))]
dat2 <- dat2[order(dat2$ID),]
identical(as.character(colnames(ex2)), as.character(dat2$ID)) # TRUE
ex.gsva2 <- ex.gsva

# remove rownames
ex.gsva <- as.data.frame(ex.gsva)
ex.gsva$gene <- rownames(ex.gsva)
rownames(ex.gsva) <- NULL

ex.gsva <- ex.gsva[,c(ncol(ex.gsva),1:ncol(ex.gsva)-1)]
head(ex.gsva)

write.table(ex.gsva, file="ex_recurrent_gsea.txt", 
  sep="\t", quote=F, col.names=T, row.names=F)

# pheno data
table(dat2$Combined)
dat2$Combined2 <- ifelse(dat2$Combined=="High Angio - High AP", "High-High", "Low-Low")
pheno <- dat2$Combined2
pheno2 <- as.data.frame(t(pheno))
data.class(pheno2)
dim(pheno2)

write.table(pheno2, file="pheno_recurrent_gsea.txt",   
  sep="\t", quote=F, col.names=F, row.names=F)
```


GSEA with clusterProfiler
```{r}
## perform Differential expression
library(ReactomePA); library(biomaRt); library(annotate); library(clusterProfiler); 
library(limma); library(enrichplot); 

# groups
group <- factor(dat2$Combined2, levels=c("Low-Low", "High-High"))
design <- model.matrix(~group)
identical(as.character(colnames(ex.gsva2)), as.character(dat2$ID))
fit <- lmFit(ex.gsva2, design)

# bayesian comparison
fit <- eBayes(fit, trend=TRUE)
de_genes <- topTable(fit, adjust="fdr", number = nrow(as.data.frame(fit)))
head(de_genes)
de_genes <- de_genes[order(de_genes$logFC, decreasing=T),] # order by Log2
de_filt <- de_genes[de_genes$adj.P.Val<=0.05 & abs(de_genes$logFC)>=1.5,] # 1156
dim(de_filt); head(de_filt)

# gene list
geneList = as.numeric(de_genes$logFC)
names(geneList) = rownames(de_genes)
geneList <- geneList[order(geneList, decreasing=T)] # enriched in High

# load genesets
load("MSigDB_broad_Allgenesets_2020.Rdata")

## calculate GSEA hallmakrs
gsea.high.hallmarks <- GSEA(geneList, TERM2GENE = hallmarks)
head(as.data.frame(gsea.high.hallmarks))
gsea.high.hallmarks$ID
write.table(as.data.frame(gsea.high.hallmarks), file="res/gsea_clusterprofiler_list_HH.txt", col.names=T, row.names=F, quote=F, sep="\t")

# plot recurrent high enriched pathways
png("res/GSEAPlot_clusterprofiles_recurrent_ALL_HH.tiff", 
     height=13, width=14, res = 600, units="in")   
gseaplot2(gsea.high.hallmarks, geneSetID = c(3,4,5,7,8,9,10,11,24,29),
  pvalue_table = TRUE)
dev.off()

# plot recurrent high enriched pathways individually
for (i in c(3,4,5,6,7,9,10,11,15,24,29)){
  anno <- gsea.high.hallmarks[i, c("NES", "pvalue", "p.adjust")]
  lab <- paste0(names(anno), "=",  round(anno, 3), collapse="\n")

  png(paste0("res/GSEAPlot_clusterprofiles_recurrent_ALL_HH_", rownames(anno), ".tiff"), 
       height=4.25, width=5, res = 600, units="in")   
    gseaplot2(gsea.high.hallmarks, geneSetID = i, 
    title = gsea.high.hallmarks$Description[i]) +
    annotate("text", 0.7, 0.75, label = lab, hjust=0, vjust=0, size=4.5)
  dev.off()
}
``` 

Functional analysis with enrichR and ReactomePA
```{r}
### Geneset enrichment reactomepa
library(ReactomePA); library(enrichplot); library(DOSE)

## being ex_nes a preranked expression matrix ordered manually by any differential expression we want to do with limma or others
#geneset enrichment

annot <- read.delim("/mnt/hydra/ubs/shared/users/Sandra/annot/human_annot/data/annot_hgcn_genes.txt", sep="\t", header=T, stringsAsFactors=F)
de_genes_annot <- merge(de_genes, annot[,c("entrez_id", "symbol")], by.x="row.names", by.y="symbol")
de_genes_annot <- de_genes_annot[order(de_genes_annot$logFC, decreasing=TRUE),]
head(de_genes_annot)
#geneList = as.numeric(de_genes_annot$logFC)
#names(geneList) = 
geneList = de_genes_annot$entrez_id

x <- enrichPathway(gene=geneList,pvalueCutoff=0.05, readable=T)
write.table(x, file="res/reactomepa_list_HH_LL.txt", col.names=T, row.names=F, quote=F, sep="\t")


png("res/geneset_enrich_HH_LL_dotplot1.tiff", 
   height=4.5, width=7, res = 600, units="in")   
  dotplot(x, showCategory=15)
dev.off()
```


## Hierarchical clustering and heatmap
```{r}
# dendograms
d1 <- dist(data.st, method = "euclidean") 
hcl1 <- hclust(d1, method="average") # samples

# rotate function
rotate.dendrogram2 <- function(x, order, ...)
  {
    if(missing(order)) { # if order is missing - return the same tree.
      warning("'order' parameter is missing, returning the tree as it was.")
      return(x)  
    }
    
    labels_x <- labels(x) 
    order_x <- order.dendrogram(x)
    number_of_leaves <- length(order_x)
    
    
    if(!is.numeric(order)) {
      order <- as.character(order)
      if(length(intersect(order, labels_x)) != number_of_leaves) {
        stop("'order' is neither numeric nor a vector with ALL of the labels (in the order you want them to be)")
      }
      order <- match(order, labels_x)
    }
    
    weights <- seq_len(number_of_leaves)
    weights_for_order <- numeric(number_of_leaves)
    weights_for_order[order_x[order]] <- weights
    reorder(x, weights_for_order, mean,...)
  }
rotate <- function(...) as.hclust(rotate.dendrogram2(as.dendrogram(hcl1),20))

# colors 
bk <- c(seq(min(as.numeric(data.st)),max(as.numeric(data.st)), length=100))

# Data frame with column annotations
dat$Recurrence <- ifelse(dat$DFS==1, "Recurrent", "Non-recurrent")
table(dat$Recurrence)
table(dat$Combined)

mat_col <- data.frame('Recurrence' = dat$Recurrence,
  "Dataset" = dat$dataset,
  "Combined" = dat$Combined,
  "Chr3_status" = dat$chr3_status)
rownames(mat_col) <- dat$ID

# List with colors for each annotation.

#colors
library(RColorBrewer); library(gplots)
ncol <- length(unique(dat$dataset))
cols_palette1 <- brewer.pal(8, "Dark2")[c(1:4,6:8)]
cols_palette2 <- brewer.pal(8, "RdBu")[c(1:4,6:8)]
cols_palette <- c(cols_palette1, cols_palette2)[1:ncol]


ncol1 <- length(unique(dat$dataset))
ncol2 <- length(unique(dat$chr3_status))
cols_palette1 <- brewer.pal(8, "Dark2")[c(1:4,6:8)][1:ncol1]
cols_palette2 <- brewer.pal(8, "RdBu")[c(1,3,7)]


mat_colors <- list('Recurrence' =  c("grey90", "grey40"),
  "Dataset"=cols_palette1, 
  "Combined"= c("firebrick3", "lightpink1","skyblue1", "dodgerblue4"),
  "Chr3_status" = c("#4393C3", "#F4A582", "grey60", "#B2182B"))
names(mat_colors$'Recurrence') <- as.character(names(table(dat$Recurrence)))
names(mat_colors$'Dataset') <- as.character(names(table(dat$dataset)))
names(mat_colors$'Combined') <- as.character(names(table(dat$Combined)))
names(mat_colors$'Chr3_status') <- as.character(names(table(dat$chr3_status)))


library(pheatmap)
heatmap <- pheatmap(mat = t(data.st),
         color = colorpanel(n=length(bk)-1, low="blue3", mid="snow", high="red3"),
         border_color = NA, show_colnames = FALSE, 
         show_rownames = TRUE,
         cluster_cols = rotate(hcl1,3:1), 
         annotation_col = mat_col, 
         annotation_colors = mat_colors,
         #drop_levels = TRUE, 
         fontsize = 7, cellwidth=1.5, cellheight=15)

heatmap

## euclidean with average and rotate 3:1

# plot heatmap
png("res/heatmap_angio-antigenpresentation.tiff", 
 height=7, width=8,  res = 600, units="in")      
heatmap
dev.off()
```

