---
title: "Code for pool analysis"
author: "García-Mulero S"
date: "12/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Meta-analysis

Load packages
```{r cars}
require(reshape); require(ggplot2); library(ggpubr); library(ggthemes); 
library(survival); library(survminer)
library(meta);  library(meta)
```

## Load all data and join

# TCGA dataset
```{r}
load("data/uveal_tcga_2020.Rdata")
exTCGA <- ex
infoClin$Recurrence <- ifelse(infoClin$DFS==1, "Recurrent", "Non-recurrent")
clinTCGA <- infoClin
clinTCGA$dataset <- "TCGA"
clinTCGA$ID <- clinTCGA$id

ips_tcga <- read.table("data/IPS_TCGA.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
mcp_tcga <- read.table("data/MCPcounter_TCGA.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
est_tcga <- read.table("data/ESTIMATE_TCGA.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
quantiseq_tcga <- read.table("data/quantiseq_TCGA.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
```

# GSE27831
```{r}
load("data/uveal_GSE27831_def.Rdata")
ex1 <- ex; clin1 <- clin
clin1$annot <- "GPL570"
clin1$dataset <- "GSE27831"

ips1 <- read.table("data/GSE27831/IPS.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
mcp1 <- read.table("data/GSE27831/MCPcounter.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
est1 <- read.table("data/GSE27831/ESTIMATE.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
quantiseq1 <- read.table("data/GSE27831/quantiseq.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
```

# GSE22138
```{r}
load("data/uveal_GSE22138_def.Rdata")
ex2 <- ex; 
clin2 <- clin
clin2$annot <- "GPL570"
clin2$dataset <- "GSE22138"

ips2 <- read.table("data/GSE22138/IPS.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
mcp2 <- read.table("data/GSE22138/MCPcounter.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
est2 <- read.table("data/GSE22138/ESTIMATE.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
quantiseq2 <- read.table("data/GSE22138/quantiseq.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
```

# GSE84976
```{r}
load("data/uveal_GSE84976_def.Rdata")
ex3 <- ex; ex3 <- as.data.frame(ex3)
clin3 <- clin
clin3$annot <- "GPL10558"
clin3$dataset <- "GSE84976"
clin3$DFS <- clin3$OS
clin3$DFS.Months <- clin3$OS.Months

ips3 <- read.table("data/GSE84976/IPS.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
mcp3 <- read.table("data/GSE84976/MCPcounter.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
est3 <- read.table("data/GSE84976/ESTIMATE.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
quantiseq3 <- read.table("data/GSE84976/quantiseq.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
```

# GSE73652
```{r}
load("data/uveal_GSE73652_def.Rdata")
ex4 <- ex; ex4 <- as.data.frame(ex4)
clin4 <- clin
clin4$annot <- "GPL10558"
clin4$dataset <- "GSE73652"
clin4$DFS.Months <- NA

ips4 <- read.table("data/GSE73652/IPS.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
mcp4 <- read.table("data/GSE73652/MCPcounter.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
est4 <- read.table("data/GSE73652/ESTIMATE.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
quantiseq4 <- read.table("data/GSE73652/quantiseq.txt", sep="\t", stringsAsFactors=FALSE, header=TRUE)
```

# join all clinical
```{r}
clin <- rbind(clinTCGA[,c(107, 95:96, 106)],
  clin1[, c(1, 19:20, 22)],
  clin2[, c(16, 17:18, 20)],
  clin3[, c(7, 12:13, 11)],
  clin4[, c(9, 10, 13, 12)]) ## 213 samples
dat <- clin
```


## Function Meta-analysis Cox (several studies, one score)
```{r}
#########################################
meta_calculate <- function(dat, score){
  meta.df <- list()
  for (i in unique(dat$dataset)){ # for each dataset
    dati <- dat[dat$dataset==i,] # take values of only one dataset
    meta <- data.frame("Study"=c(NA), "HR"=c(NA), "error.HR" = c(NA)) # create empty df
    cox <- coxph(Surv(DFS.Months, DFS) ~ dati[,score], data=dati) # calculate cox
    meta$Study[1] <- i # add info of cox results for this study
    meta$HR[1] <-summary(cox)$coefficients[2]
    meta$error.HR[1] <- summary(cox)$coefficients[3]
    meta.df[[i]] <- meta
  }
  meta.df <- do.call(rbind, meta.df) # joint all results in one df
  return(meta.df)
}
#########################################

```


## Function Pool analysis Cox (several scores)
```{r}
#########################################
pool_calculate <- function(dat){
  meta.df <- list()
  for (i in 5:ncol(dat)){ # from column 5 are all the scores (can change)
    meta <- data.frame("Score"=c(NA), "HR"=c(NA), "error.HR" = c(NA)) # create df
    cox <- coxph(Surv(DFS.Months, DFS) ~ dat[,i] + strata(dataset), data=dat) # cox
    meta$Score[1] <- colnames(dat)[i] # add all data for this score
    meta$HR[1] <-summary(cox)$coefficients[2]
    meta$error.HR[1] <- summary(cox)$coefficients[3]
    meta.df[[i]] <- meta
  }
  meta.df <- do.call(rbind, meta.df)
  return(meta.df)
}
#########################################

```


## Meta-analysis IPS 
```{r}
ips <- cbind(ips_tcga, ips1, ips2, ips3)
ips <- t(ips)
rownames(ips) <- gsub("\\.", "-", rownames(ips))
ips <- 100*(ips)

# merge
dat_ips <- merge(dat, ips, by.x="ID", by.y="row.names")
dat_ips <- dat_ips[,c(1:9)]
colnames(dat_ips)[5:9] <- c("Antigen presentation", "Effector cells", "Supressor cells", "Checkpoints", "Aggregated Score")

## Meta-analysis dataset per dataset for each score
meta.df1 <- meta_calculate(dat=dat_ips, score="Antigen presentation")
meta.res1 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df1, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res1, digits=2)

jpeg("res/uveal_meta-analysis_all_IPS_AP.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res1,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     smlab = "",
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Antigen presentation")), .5, .9, gp = gpar(fontsize=8))

dev.off()
   

meta.df2 <- meta_calculate(dat=dat_ips, score="Effector cells")
meta.res2 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df2, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res2, digits=2)

jpeg("res/uveal_meta-analysis_all_IPS_EC.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res2,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Effector cells")), .5, .9, gp = gpar(fontsize=8))

dev.off()

meta.df3 <- meta_calculate(dat=dat_ips, score="Supressor cells")
meta.res3 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df3, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res3, digits=2)

jpeg("res/uveal_meta-analysis_all_IPS_SC.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res3,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Supressor cells")), .5, .9, gp = gpar(fontsize=8))

dev.off()

meta.df4 <- meta_calculate(dat=dat_ips, score="Aggregated Score")
meta.res4 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df4, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res4, digits=2)

jpeg("res/uveal_meta-analysis_all_IPS_Aggregated.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res4,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Aggregated Immunophenoscore")), .5, .9, gp = gpar(fontsize=8))

dev.off()

meta.df5 <- meta_calculate(dat=dat_ips, score="Checkpoints")
meta.res5 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df5, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res5, digits=2)

jpeg("res/uveal_meta-analysis_all_IPS_Checkpoints.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res5,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Checkpoints")), .5, .9, gp = gpar(fontsize=8))
dev.off()
```


## Pool analysis IPS: Study stratified Cox  
```{r}
dat_ips$DFS <- as.numeric(dat_ips$DFS);
dat_ips$DFS.Months <- as.numeric(dat_ips$DFS.Months)
meta.df <- pool_calculate(dat_ips)

## Meta-analysis
meta.df <- meta.df[order(meta.df$HR, decreasing=F),]
rownames(meta.df) <- NULL
meta.res <- metagen(TE=log(HR), seTE=error.HR, data=meta.df, studlab=Score, sm = "HR")
print(meta.res, digits=2)

# plot
jpeg("res/uveal_pool_analisis_all_IPS.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res,
     sortvar=TE, 
     comb.fixed=FALSE,
     overall = FALSE,
     pooled.totals = TRUE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Score","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=TRUE)

dev.off()
```


## Meta-analysis ESTIMATE 
```{r}
est <- cbind(est_tcga, est1, est2, est3)
est["StromalScore",]  <- (est["StromalScore",])/100
est["ImmuneScore",]  <- (est["ImmuneScore",])/100
est["TumorPurity",]  <- 100*(est["TumorPurity",])
est["ESTIMATEScore",]  <- (est["ESTIMATEScore",])/100
summary(melt(est)$value)

est <- t(est)
rownames(est) <- gsub("\\.", "-", rownames(est))

# merge
dat_est <- merge(dat, est, by.x="ID", by.y="row.names")

##  Meta-analysis dataset per dataset of each score
meta.df1 <- meta_calculate(dat=dat_est, score="StromalScore")
meta.res1 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df1, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res1, digits=2)

jpeg("res/uveal_meta-analysis_all_EST_StromalScore.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res1,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Stromal Score")), .5, .9, gp = gpar(fontsize=8))

dev.off()

meta.df2 <- meta_calculate(dat=dat_est, score="ImmuneScore")
meta.res2 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df2, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res2, digits=2)

jpeg("res/uveal_meta-analysis_all_EST_ImmuneScore.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res2,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Immune Score")), .5, .9, gp = gpar(fontsize=8))

dev.off()


meta.df3 <- meta_calculate(dat=dat_est, score="ESTIMATEScore")
meta.res3 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df3, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res3, digits=2)

jpeg("res/uveal_meta-analysis_all_EST_ESTIMATEScore.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res3,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("ESTIMATE Score")), .5, .9, gp = gpar(fontsize=8))

dev.off()


meta.df4 <- meta_calculate(dat=dat_est, score="TumorPurity")
meta.res4 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df4, sm = "HR", studlab=Study, 
  comb.fixed=FALSE, title="tumorpuritytest", keepdata=F)
print(meta.res4, digits=2)

jpeg("res/uveal_meta-analysis_all_EST_TumorPurity.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res4,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Tumor Purity")), .5, .9, gp = gpar(fontsize=8))

dev.off()
```


## Pool analysis ESTIMATE: study stratified Cox  
```{r}
dat_est$DFS <- as.numeric(dat_est$DFS);
dat_est$DFS.Months <- as.numeric(dat_est$DFS.Months)
meta.df <- pool_calculate(dat_est)

## Meta-analysis
meta.df <- meta.df[order(meta.df$HR, decreasing=F),]
rownames(meta.df) <- NULL
meta.res <- metagen(TE=log(HR), seTE=error.HR, data=meta.df, studlab=Score, sm = "HR")
print(meta.res, digits=2)

# plot
jpeg("res/uveal_pool_analysis_all_EST.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res,
     sortvar=TE, 
     comb.fixed=FALSE,
     overall = FALSE,
     pooled.totals = TRUE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Score","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=TRUE)

dev.off()
```


## Pool analysis MCP-counter 
```{r}
mcp <- cbind(mcp_tcga, mcp1, mcp2, mcp3)
summary(melt(mcp)$value)

# merge
dat_mcp <- merge(dat, t(mcp), by.x="ID", by.y="row.names")
dat_mcp <- dat_mcp[, colnames(dat_mcp)!="CD8 T cells"]

##  Meta-analysis dataset per dataset of interesting cell types

# B cells
meta.df1 <- meta_calculate(dat=dat_mcp, score="B lineage")
meta.res1 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df1, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res1, digits=2)

jpeg("res/uveal_meta-analysis_all_mcpcounter_Bcells.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res1,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("B cells")), .5, .9, gp = gpar(fontsize=8))

dev.off()


## Study stratified Cox  
dat_mcp$DFS <- as.numeric(dat_mcp$DFS);
dat_mcp$DFS.Months <- as.numeric(dat_mcp$DFS.Months)
meta.df <- pool_calculate(dat_mcp)
meta.df <- meta.df[order(meta.df$HR, decreasing=F),]
rownames(meta.df) <- NULL
meta.df

## Meta-analysis
meta.res <- metagen(TE=log(HR), seTE=error.HR, data=meta.df, studlab=Score, sm = "HR")
print(meta.res, digits=2)

# plot
jpeg("res/uveal_pool_analisis_all_MCP-counter.jpg", 
  bg = "white", res = 600, width = 5, height = 3, units = "in")
forest(meta.res,
     sortvar=TE, 
     comb.fixed=FALSE,
     overall = FALSE,
     pooled.totals = TRUE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Score","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=TRUE)

dev.off()
```


## Pool  analysis Quantiseq 
```{r}
quantiseq <- cbind(quantiseq_tcga, quantiseq1, quantiseq2, quantiseq3)
summary(melt(quantiseq)$value)
quantiseq <- 100*(quantiseq)
colnames(quantiseq) <- gsub("\\.", "-", colnames(quantiseq))

# merge
dat_quantiseq <- merge(dat, t(quantiseq), by.x="ID", by.y="row.names")
head(dat_quantiseq)


##  Meta-analysis dataset per dataset of interesting cell types

# B cells
meta.df1 <- meta_calculate(dat=dat_quantiseq, score="B cell")
meta.res1 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df1, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res1, digits=2)

jpeg("res/uveal_meta-analysis_all_quantiseq_Bcells.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res1,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("B cells")), .5, .9, gp = gpar(fontsize=8))

dev.off()

# Macrophage M2
meta.df2 <- meta_calculate(dat=dat_quantiseq, score="Macrophage M2")
meta.res2 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df2, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res2, digits=2)

jpeg("res/uveal_meta-analysis_all_quantiseq_MacrophageM2.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res2,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("Macrophage M2")), .5, .9, gp = gpar(fontsize=8))

dev.off()


# NK cell
meta.df3 <- meta_calculate(dat=dat_quantiseq, score="NK cell")
meta.res3 <- metagen(TE=log(HR), seTE=error.HR, data=meta.df3, sm = "HR", studlab=Study, comb.fixed=FALSE)
print(meta.res3, digits=2)

jpeg("res/uveal_meta-analysis_all_quantiseq_NKcell.jpg", 
  bg = "white", res = 600, width = 4, height = 2, units = "in")
forest(meta.res3,
     sortvar=TE, 
     comb.fixed=FALSE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     test.overall.random = TRUE,
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Study","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=FALSE)
grid.text(expression(bold("NK cells")), .5, .9, gp = gpar(fontsize=8))

dev.off()

## Study stratified Cox  
dat_quantiseq$DFS <- as.numeric(dat_quantiseq$DFS);
dat_quantiseq$DFS.Months <- as.numeric(dat_quantiseq$DFS.Months)
meta.df <- pool_calculate(dat_quantiseq)
meta.df <- meta.df[order(meta.df$HR, decreasing=F),]
meta.df

## Meta-analysis
meta.res <- metagen(TE=log(HR), seTE=error.HR, data=meta.df, studlab=Score, sm = "HR")
print(meta.res, digits=2)

# plot
jpeg("res/uveal_pool_analisis_all_quantiseq.jpg", 
  bg = "white", res = 600, width = 5, height = 3, units = "in")
forest(meta.res,
     sortvar=TE, 
     comb.fixed=FALSE,
     overall = FALSE,
     pooled.totals = TRUE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Score","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=TRUE)

dev.off()
```


## Pool analysis consensusTME
```{r}
# with all signatures
consensusTME <- ConsensusTME::consensusGeneSets$Unfiltered

# perfomr GSVA function
library(GSVA)
GSVA.TCGA <- gsva(expr=as.matrix(exTCGA), gset.idx.list=consensusTME)
GSVA1 <- gsva(expr=as.matrix(ex1), gset.idx.list=consensusTME)
GSVA2 <- gsva(expr=as.matrix(ex2), gset.idx.list=consensusTME)
GSVA3 <- gsva(expr=as.matrix(ex3), gset.idx.list=consensusTME)
GSVA4 <- gsva(expr=as.matrix(ex4), gset.idx.list=consensusTME)

# join GSVA
GSVA <- cbind(GSVA.TCGA, GSVA1, GSVA2, GSVA3)

# add to data
summary(melt(GSVA)$value)
consensus <- 100*(GSVA)

# merge
dat_consensus <- merge(dat, t(consensus), by.x="ID", by.y="row.names")
head(dat_consensus)

# pool analysis
dat_consensus$DFS <- as.numeric(dat_consensus$DFS);
dat_consensus$DFS.Months <- as.numeric(dat_consensus$DFS.Months)
meta.df <- pool_calculate(dat_consensus)
meta.df <- meta.df[order(meta.df$HR, decreasing=F),]
meta.df

# Meta-analysis
meta.res <- metagen(TE=log(HR), seTE=error.HR, data=meta.df, studlab=Score, sm = "HR")
print(meta.res, digits=2)

# plot
jpeg("res/uveal_pool_analisis_all_consensusTME.jpg", 
  bg = "white", res = 600, width = 5, height = 4, units = "in")
forest(meta.res,
     sortvar=TE, 
     comb.fixed=FALSE,
     overall = FALSE,
     pooled.totals = TRUE,
     rightcols=c("w.random"),
     rightlabs = c("Weight"),
     leftcols=c("studlab", "effect", "ci"),
     leftlabs=c("Score","HR","95% CI"),
     text.random = "Overall effect",
     print.tau2 = FALSE,
     col.diamond.lines = "black",
     col.predict = "black",
     print.I2.ci = FALSE,
     digits.sd = 2,
     fontsize= 8,
     plotwidth="1.5inch",
     just="left",
     spacing=.8,
     addrow=FALSE, addrow.overall=TRUE)

dev.off()
```

